DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report

### 2022-10-06 Draft Ped Metrics Example.Rmd
add weighted scores for summary

```{r combine-rd-sw-sh, echo = TRUE}
# combine the sidewalk and shoulder and roads data
comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
                                     SEGMENT_ID_R = SEGMENT_ID,
                                     FULL_NAME_R = FULL_NAME, 
                                     LENGTH_R = LENGTH,
                                     road_class, RUCODE), 
                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
  mutate(distance = as.integer(st_length(.)))
#distance is the length of the matched road segment and sidewalk or shoulder 
comb %>% st_set_geometry(., NULL) %>% 
  group_by(road_class, SIDE_OF_RO, TYPE) %>%
  summarize(s_miles = sum(distance)/5280,
            miles = format(round(as.numeric(sum(distance)/5280), 1), 
                           digits = 1, nsmall=1, big.mark=","),
            .groups = "drop")
```

@@

knitr::kable(road_standards,
             caption = "Simplified Assumptions to Allow for Evaluation]")  %>%
   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation")
```

@@

Interpretation of Scores for Segments and for Areas

|Score| Interpretation |
|------|:-----|
|10|Good|
|5 TO 9 | Adequate^[For a segment, this means one side is good or both sides meet AASHTO minimums]|
|3 AND 4 | Minimal^[For a segment, this means one side meets AASHTO minimums]|
|Less than 3 |Poor|

## Results

@@

road_standards <- road_standards %>%
  rename(standard_sidewalk = sidewalk, 
         standard_shoulder = shoulder, 
         road_class = class)
  # pivot_longer(cols=c(sidewalk, shoulder), 
  #              names_to = "Standard Pedestrian Facility",
  #              values_to = "Standard Width") %>% filter
sw_ncdf <- sidewalks_nc %>%
  mutate(sidewalk_right = ifelse(SIDE_OF_RO == "RIGHT", WIDTH, NA),
         sidewalk_left = ifelse(SIDE_OF_RO == "LEFT", WIDTH, NA)) %>%
  select(RD_LOG_ID, SEGMENT_ID, sidewalk_right, sidewalk_left)
sh_ncdf <- shoulders_nc %>%
  mutate(shoulder_right = ifelse(SIDE_OF_RO == "RIGHT", WIDTH, NA),
         shoulder_left = ifelse(SIDE_OF_RO == "LEFT", WIDTH, NA)) %>%
  select(RD_LOG_ID, SEGMENT_ID, shoulder_right, shoulder_left)
sw_sh <- full_join(st_set_geometry(sw_ncdf, NULL), 
                   st_set_geometry(sh_ncdf, NULL)
                   )

@@

comb_stand <- comb %>% 
  st_set_geometry(., NULL) %>% 
  mutate(setting = ifelse(RUCODE == 1, "Rural", 
                          ifelse(RUCODE == 2, "Urban", "Unknown")),
         ped_type = sub("\\:.*", "", TYPE),
         #ped_w = as.numeric(sub(".*?([0-9]+).*", "\\1", TYPE)),
         nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
  #select(RD_LOG_ID = RD_LOG_ID_R, SEGMENT_ID = SEGMENT_ID_R, FULL_NAME = FULL_NAME_R,
  #       SEGMENT_LENGTH = LENGTH_R, SW_SH_LENGTH = LENGTH, SIDE_OF_RO,
  #       distance, road_class, setting, nn, ped_w) %>%
  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), #c(RD_LOG_ID, SEGMENT_ID, FULL_NAME), 
              names_from = nn, 
              values_from = WIDTH,
              values_fn = mean) %>%
  left_join(road_standards)

@@

knitr::kable(
  comb_score %>%
  group_by(setting, road_class, score) %>% 
  summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
            segments = nice_comma(n()), 
            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
            .groups = "drop") 
  )
```

>>>>

```{r combine-rd-sw-sh, echo = TRUE}
# combine the sidewalk and shoulder and roads data
x <- st_as_sf(nc_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
x$RD_LOG_ID <- 43809
comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
                                     SEGMENT_ID_R = SEGMENT_ID,
                                     FULL_NAME_R = FULL_NAME, 
                                     LENGTH_R = LENGTH,
                                     road_class, RUCODE), 
                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
  mutate(distance = as.integer(st_length(.)))
#distance is the length of the matched road segment and sidewalk or shoulder 
comb %>% st_set_geometry(., NULL) %>% 
  group_by(road_class, SIDE_OF_RO, TYPE) %>%
  summarize(s_miles = sum(distance)/5280,
            miles = format(round(as.numeric(sum(distance)/5280), 1), 
                           digits = 1, nsmall=1, big.mark=","),
            .groups = "drop")
```

@@

knitr::kable(road_standards,
             caption = "Simplified Assumptions to Allow for Evaluation")  %>%
   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation")
```

@@

```{r}
score_interpretation <- tribble(
  ~score_at_least, ~interpretation,
  10, "Good",
  5, "Adequate",
  3, "Minimum",
  0, "Poor",
)

getScoreInterpretation <- function(x){
  first(score_interpretation$interpretation[
        score_interpretation$score_at_least <= x]
    )
}

knitr::kable(score_interpretation,
             caption = "Interpretation of Scores for Segments and for Areas")  %>%
   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums")
```

## Results
First, we calculate the score overall for the county and then break it down for other smaller units of area in our community.

@@

road_standards <- road_standards %>%
  rename(standard_sidewalk = sidewalk, 
         standard_shoulder = shoulder, 
         road_class = class)

@@

comb_stand <- comb %>% 
  st_set_geometry(., NULL) %>% 
  mutate(setting = ifelse(RUCODE == 1, "Rural", 
                          ifelse(RUCODE == 2, "Urban", "Unknown")),
         ped_type = sub("\\:.*", "", TYPE),
         nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), 
              names_from = nn, 
              values_from = WIDTH,
              values_fn = mean) %>%
  left_join(road_standards)

@@

```{r}
comb_score2 <- st_set_geometry(ncdf, NULL) %>% 
  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
  mutate(setting = ifelse(RUCODE == 1, "Rural", 
                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
  left_join(comb_score) %>%
  mutate(the_score = ifelse(is.na(score), 0, score),
         the_length = ifelse(is.na(distance), LENGTH, distance),
         w_score = the_score*the_length)

comb_total_score <- sum(comb_score2$w_score)/sum(comb_score2$the_length)

comb_score_by_setting <- comb_score2 %>%
  group_by(setting) %>%
  summarise(score = sum(w_score)/sum(the_length),
            .groups = "drop") %>%
  rowwise() %>%
  mutate(interpretation = getScoreInterpretation(score))

comb_score_by_setting_class <- comb_score2 %>%
  group_by(setting, road_class) %>%
  summarise(score = sum(w_score)/sum(the_length),
            .groups = "drop") %>%
  rowwise() %>%
  mutate(interpretation = getScoreInterpretation(score))
```

The overall score for pedestrian facilities in Kitsap County is `r nice_comma(comb_total_score)`, `r getScoreInterpretation(comb_total_score)`.

```{r}
knitr::kable(comb_score_by_setting_class)
```

More detailed information
```{r}
knitr::kable(
  comb_score2 %>%
  group_by(setting, road_class, score) %>% 
  summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
            segments = nice_comma(n()), 
            #road_len = nice_comma(sum(the_len, na.rm=TRUE)/5280, 1),
            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
            .groups = "drop") 
  )
```

```{r}
ggplot(filter(comb_score2, setting != "Unknown"), 
       aes(x = as.factor(score), y = the_length/5280))+
  geom_col() +
  labs(x = "score based on data available", 
       y = "centerline miles of road") +
  theme_minimal() +
  facet_grid(setting ~ road_class)
```

If we remove the segments where we do not have information about sidewalks and shoulders at all, we can see that for those, we are generally finding higher scores for rural roads - with paved shoulders.

```{r}
ggplot(filter(comb_score2, setting != "Unknown", !is.na(score)), 
       aes(x = as.factor(score), y = LENGTH/5280))+
  geom_col() +
  labs(x = "score based on data available", 
       y = "centerline miles of road") +
  theme_minimal() +
  facet_grid(setting ~ road_class)
```

Some of the segments with score of 0 are because they have shoulders when they should have sidewalks. Here's a few example segments. If you are familiar with Central Kitsap, you might recognize this on Provost - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate.

```{r}

roads_0 <- comb_score2 %>%
  filter(score == 0) %>%
  select(setting, road_class, FULL_NAME_R, sidewalk_right:score)

knitr::kable(head(roads_0))

```{r}
roads_10 <- comb_score2 %>%
  filter(score == 10) %>%
  group_by(setting, road_class, "Road Name" = FULL_NAME_R) %>%
  summarise(segments = nice_comma(n()), 
            road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
            .groups = "drop")

knitr::kable(roads_10)
```

### 2022-10-06 child_single_geo.Rmd
add weighted scores for summary

>>>>

```{r eval=has_both}
this_comb_stand <- this_comb %>% 
  st_set_geometry(., NULL) %>% 
  mutate(setting = ifelse(RUCODE == 1, "Rural", 
                          ifelse(RUCODE == 2, "Urban", "Unknown")),
         ped_type = sub("\\:.*", "", TYPE),
         nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), 
              names_from = nn, 
              values_from = WIDTH,
              values_fn = mean) %>%
  left_join(road_standards)

this_comb_score <- this_comb_stand %>%
  mutate(score_left = getSideScore(standard_sidewalk, standard_shoulder, 
                                   sidewalk_left, shoulder_left),
         score_right = getSideScore(standard_sidewalk, standard_shoulder, 
                                   sidewalk_right, shoulder_right),
         score = score_left + score_right)
this_comb_score2 <- st_set_geometry(this_roads, NULL) %>% 
  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
  mutate(setting = ifelse(RUCODE == 1, "Rural", 
                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
  left_join(this_comb_score) %>%
  mutate(the_score = ifelse(is.na(score), 0, score),
         the_length = ifelse(is.na(distance), LENGTH, distance),
         w_score = the_score*the_length)

this_comb_total_score <- sum(this_comb_score2$w_score)/sum(this_comb_score2$the_length)
```
